---
title: "Topic 2.2  Exercise 6"
author: "Princess Mae Bigar"
date: "2025-09-01"
output: html_document
---
## **Analyzing Gene Expression Data**

```{r, comments=''}

#Creating the data frame
gene_expression <- data.frame(
  Gene = c("GeneA", "GeneB", "GeneC", "GeneD",
           "GeneE", "GeneA", "GeneB", "GeneC",
           "GeneD", "GeneE", "GeneA", "GeneB",
           "GeneC", "GeneD", "GeneE"),
  Expression_Level = c(15.3,10.8, 12.5, 18.1,
                       11.7, 20.2, 17.6, 14.3,
                       19.5, 16.2,13.8, 21.1,
                       19.7, 22.5, 20.3),
  Tissue = c("Tissue1", "Tissue1", "Tissue1",
             "Tissue1", "Tissue1", "Tissue2",
             "Tissue2", "Tissue2", "Tissue2",
             "Tissue2", "Tissue3", "Tissue3",
             "Tissue3", "Tissue3", "Tissue3")
)

#Display the data frame
print("Gene Expression Data:")
print(gene_expression)

#Using for loop and if statement to find genes with expression > 20
print("Genes with Expression Level above 20:")
for (i in 1:nrow(gene_expression)) {
  if (gene_expression$Expression_Level[i] > 20) {
    print(paste(gene_expression$Gene[i],
                gene_expression$Tissue[i],
 gene_expression$Expression_Level[i]))
  }
}

#Using while loop to filter Tissue2 data until finding first gene with expression < 15
tissue2_data <- gene_expression[gene_expression$Tissue == "Tissue2", ]
row_index <- 1
found <- FALSE

print("Checking Tissue2 data for first gene with expression < 15:")

while (row_index <= nrow(tissue2_data)) {
  current_gene <- tissue2_data[row_index, ]
  if (current_gene$Expression_Level < 15) {
    print(paste("First gene found with expression level below 15:", current_gene$Gene, "- Tissue:", current_gene$Tissue, "- Expression:", current_gene$Expression_Level))

    found <- TRUE
    break
  }
  row_index <- row_index + 1
}

```

```{r, comments=''}
## **Monitoring Specie Populations**

#Creating the species population vector
species_populations <- c(150, 45, 75, 200, 30)

print("Initial species populations:")
print(species_populations)

#Use for loop and if-else to classify species status
print("Species conservation status classification:")

for (i in 1:length(species_populations)) {
  if (species_populations[i] < 50) {
    print(paste("Species", i, "- Population:", species_populations[i], "- Status: Endangered"))
  } else {
    print(paste("Species", i, "- Population:", species_populations[i], "- Status: Stable"))
  }
}

#Assigning 
growth_rate <- 0.05 #5% yearly growth rate
years <- c(1,2,3)

print("Population growth simulation over 3 years(5% growth rate);")

#Use nested for loops to stimulate population growth
for (species in species_populations) {
  for (year in years) {
    population <- species * (1.05)^year
    print(paste("Species population", species, "Year", year, "=", round(population, 2)))
  }
}

```

## **Monitoring Water Quality**
```{r, comments=''}

#Create pollution levels vector 
pollution_levels <- c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)

print("Pollution measurements:")
print(pollution_levels)

print("Pollution level warnings:")

for (level in pollution_levels) {
  if (level < 50) {
    next  # Skip measurements below 50
  }
  
  if (level > 50) {
    print(paste("WARNING: Pollution level", level, "ppm - Above safety threshold!"))
  }
}

#Use repeat loop to monitor until critical level of 90 ppm is detected
print("Continuous pollution monitoring (until critical level detected):")
current_index <- 1

repeat {
  current_level <- pollution_levels[current_index]
  
  print(paste("Monitoring... Current level:", current_level, "ppm"))
  
  if (current_level >= 90) {
    print(paste("CRITICAL WARNING: Pollution level", current_level, 
                "ppm detected! Immediate action required!"))
    break  # Stop the loop when critical level is reached
  }
  
  current_index <- current_index + 1
  
   if (current_index > length(pollution_levels)) {
    print("Monitoring complete - no critical levels reached")
    break
  }
}
