---
title: "Untitled"
author: "Princess Mae Bigar"
date: "2025-08-30"
output: html_document
---
## **Analyzing Nucleotide Sequence Data**

```{r, comment=''}
# Create a character vector of DNA sequences
dna_sequences <- c("ATGCGTACGTAGCTAGCGT","ATGCGTACGTAGTTAGCGT", "ATGCGTACGTAGATAGCGT")

# Calculate lengths of sequences
sequence_lengths <- nchar(dna_sequences)
print("Length of each DNA sequence:")
print(sequence_lengths)

# Count number of "A" in each sequence 
library(stringr)
a_counts <- str_count(dna_sequences, "A")
print("Number of 'A' nucleotides in each sequence:")
print(a_counts)

```

## **Monitoring Species Population in Different Habitats**

```{r, comment=''}
# Create a data frame for biodiversity data
biodiversity_data <- data.frame(
  Species = c("Oak", "Pine", "Maple","Birch", "Spruce", "Cedar","Ash", "Fir", "Beech"),
  Count = c(120, 85, 100,60, 150, 75, 55, 95, 110),
  Habitat = c("Forest A", "Forest A", "Forest A","Forest B", "Forest B", "Forest B","Forest C", "Forest C", "Forest C")
)

print(biodiversity_data)

# Calculate total number of species observed per habitat
print("Total number of species per habitat")
total_species_per_habitat <- print(tapply(biodiversity_data$Count, biodiversity_data$Habitat, sum))

# Subset species with count less than 50
low_count_species <- subset(biodiversity_data, Count < 50)

if (nrow(low_count_species) == 0) {
  print("Species with count less than 50: 0")
} else {
  print(paste("Species with count less than 50:", paste(low_count_species$Species)))
}

```

## **Monitoring Temperature Changes**

```{r, comment=''}
# Create the temperature matrix
monthly_temps <- matrix(
  c(2.5, 3.1, 6.2, 10.5, 15.3, 20.1, 22.4, 21.9, 18.3, 13.8, 7.5, 3.9, 2.7, 3.5, 6.8, 11.0, 16.0, 21.0, 23.1, 22.5, 19.0, 14.5, 8.0, 4.1),
  nrow = 12, ncol = 2, byrow = FALSE)

# Name rows and columns (remove underscores)
rownames(monthly_temps) <- c("January", "February", "March", "April", "May", "June",
                             "July", "August", "September", "October", "November", "December")
colnames(monthly_temps) <- c("Year 1", "Year 2")

print(monthly_temps)

# Calculate average temperature for each year
print("Average temperatures per year")
avg_temps <- print(colMeans(monthly_temps))

# Check if Year 2 had higher average temperature than Year 1
year2_higher <- avg_temps["Year 2"] > avg_temps["Year 1"]
print(paste("Is Year 2 higher than Year 1 on average:", year2_higher))

# Subset summer months (June, July, August)
summer_months <- monthly_temps[c("June", "July", "August"), ]
print("Summer Months (June-August) temperatures:")
print(summer_months)

